---
title: "Supplementary_Methods--Infant Urobiome"
author:   
  - Seth Reasoner
  - Viktor Flores
  - Grace Morales
  - Maria Hadjifrangiskou
date: "2023-01-7"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
---

# Introduction

This file includes documentation detailing the code used for bioinformatics and statistical analysis accompanying the manuscript **Defining the Infant Male Urobiome and Moving Towards Mechanisms in Urobiome Research**. Catheterized urine samples were collected from 50 healthy male infants. Following DNA extraction, the V4 hypervariable 16S rRNA region was amplified using the 515F and 806R primers. Sequences can be accessed under BioProject PRJNA912725. Supplementary Table S6 includes all sample accession information. Two sequencing runs were completed--one on a Illumina MiSeq and one on a Illumina NovaSeq. These runs and the samples sequenced are detailed in the manuscript text and Supplementary Table S1. Initial bioinformatic processing of 16S rRNA reads will be presented twice, once for each run. As detailed in Supplementary Table S1, **each sequencing run included 8 extraction blanks, 4 PCR negative controls, and 8 serial dilutions of a mock microbial community**. 

# Processing of V4 16S rRNA reads using dada2


## Processing of 16S rRNA reads from Illumina MiSeq run

```{r, eval=FALSE}
##Load requisite libraries (libraries used in each code chunk are re-called)
library("tidyverse")
library("dada2")
library("here")
library("phyloseq")

##We would like to credit the following tutorial for guiding this code:
##"https://ycl6.github.io/16S-Demo/index.html"

##Create Folders
MiSeq_fastq = "/MiSeq_fastq/"        # raw fastq files
MiSeq_filtered = "/MiSeq_filtered"  # dada2 trimmed fastq files
MiSeq_output = "/MiSeq_output"      # output files
MiSeq_images = "/MiSeq_images"       # output images

if(!dir.exists(MiSeq_filtered)) dir.create(MiSeq_filtered)
if(!dir.exists(MiSeq_output)) dir.create(MiSeq_output)
if(!dir.exists(MiSeq_images)) dir.create(MiSeq_images)

##Create List of File Names
fns_MiSeq = sort(list.files(MiSeq_fastq, full.names = TRUE))
fnFwds_MiSeq = fns_MiSeq[grep("R1.fastq", fns_MiSeq)]
fnRevs_MiSeq = fns_MiSeq[grep("R2.fastq", fns_MiSeq)]

MiSeq_sample_names = gsub("_R1.fastq", "", basename(fnFwds_MiSeq))

##Visualize Read Quality
ii = 1:length(MiSeq_sample_names)
pdf(paste0(MiSeq_images, "/plotQualityProfile.pdf"), width = 8, height = 8, pointsize = 12)
for(i in ii) {
  message(paste0("[", i ,"/", length(MiSeq_sample_names), "] ", MiSeq_sample_names[i]))
  print(plotQualityProfile(fnFwds_MiSeq[i]) + ggtitle("Fwd"))
  print(plotQualityProfile(fnRevs_MiSeq[i]) + ggtitle("Rev"))
}
invisible(dev.off())

##Filter and Trim
filtFwds_MiSeq = file.path(MiSeq_filtered, basename(fnFwds_MiSeq))
filtRevs_MiSeq = file.path(MiSeq_filtered, basename(fnRevs_MiSeq))

out_MiSeq = filterAndTrim(fnFwds_MiSeq, filtFwds_MiSeq, fnRevs_MiSeq, filtRevs_MiSeq,
                    truncLen = c(140,140), minLen = 120, maxN = 0, truncQ = 2, maxEE =c(2,2),
                    rm.phix = TRUE, compress = TRUE, verbose = TRUE, multithread = TRUE)

out_MiSeq = as.data.frame(out_MiSeq)
rownames(out_MiSeq) = MiSeq_sample_names
head(out_MiSeq, 10)

##Learn the Error Rates
errFwd_MiSeq = learnErrors(filtFwds_MiSeq, multithread = TRUE)
errRev_MiSeq = learnErrors(filtRevs_MiSeq, multithread = TRUE)

pdf(paste0(MiSeq_images, "/plotErrors.pdf"), width = 10, height = 10, pointsize = 12)
plotErrors(errFwd_MiSeq, nominalQ = TRUE)
plotErrors(errRev_MiSeq, nominalQ = TRUE)
invisible(dev.off())

##Sample Inference
dadaFwds_MiSeq = dada(filtFwds_MiSeq, err = errFwd_MiSeq, pool = FALSE, multithread = TRUE)
dadaRevs_MiSeq = dada(filtRevs_MiSeq, err = errRev_MiSeq, pool = FALSE, multithread = TRUE)

##Merge Paired Reads
merged_MiSeq = mergePairs(dadaFwds_MiSeq, filtFwds_MiSeq, dadaRevs_MiSeq, filtRevs_MiSeq, verbose = TRUE)

##Construct the Sequence Table
seqtab_MiSeq <- makeSequenceTable(merged_MiSeq)

table(nchar(getSequences(seqtab_MiSeq)))

##Remove Chimeras
seqtab_nochim_MiSeq <- removeBimeraDenovo(seqtab_MiSeq, method="consensus", multithread=2, verbose=TRUE)
##Percent of sequences that are non-chimeric 
sum(seqtab_nochim_MiSeq)/sum(seqtab_MiSeq)

rownames(seqtab_nochim_MiSeq) <- MiSeq_sample_names

##Track Reads through the Pipeline
getN <- function(x) sum(getUniques(x))
track_MiSeq <- cbind(out_MiSeq, sapply(dadaFwds_MiSeq, getN), sapply(dadaRevs_MiSeq, getN), sapply(merged_MiSeq, getN), rowSums(seqtab_nochim_MiSeq))
colnames(track_MiSeq) <- c("input", "filtered", "denoisedF", "denoisedR", "merged","nonchim")
rownames(track_MiSeq) <- MiSeq_sample_names
head(track_MiSeq)

##Assign Taxonomy
seqs_MiSeq = getSequences(seqtab_nochim_MiSeq)
dbpath = "/16S_DB/"
ref_db = paste0(dbpath, "silva_nr99_v138.1_train_set.fa.gz")

taxonomy_tab_MiSeq <- assignTaxonomy(seqs_MiSeq, refFasta = ref_db)

MiSeq_metadata <- read.csv("https://raw.githubusercontent.com/reaset41/Infant-Urobiome/main/MiSeq_metadata.csv",header = TRUE,row.names = 1)

##Create Phyloseq Object
MiSeq_phyloseq_unfiltered <- phyloseq(otu_table(seqtab_nochim_MiSeq, taxa_are_rows=FALSE), 
               tax_table(taxonomy_tab_MiSeq), sample_data(MiSeq_metadata))


```



## Processing of 16S rRNA reads from Illumina NovaSeq run

```{r, eval=FALSE}
##Load requisite libraries
library("tidyverse")
library("dada2")
library("here")
library("phyloseq")

##Create Folders
Nova_fastq = "/Nova_fastq/"        # raw fastq files
Nova_filtered = "/Nova_filtered"  # dada2 trimmed fastq files
Nova_output = "/Nova_output"      # output files
Nova_images = "/Nova_images"       # output images

if(!dir.exists(Nova_filtered)) dir.create(Nova_filtered)
if(!dir.exists(Nova_output)) dir.create(Nova_output)
if(!dir.exists(Nova_images)) dir.create(Nova_images)

##Create List of File Names
fns_Nova = sort(list.files(Nova_fastq, full.names = TRUE))
fnFwds_Nova = fns_Nova[grep("R1.fastq", fns_Nova)]
fnRevs_Nova = fns_Nova[grep("R2.fastq", fns_Nova)]

Nova_sample_names = gsub("_R1.fastq", "", basename(fnFwds_Nova))

##Visualize Read Quality
ii = 1:length(Nova_sample_names)
pdf(paste0(Nova_images, "/plotQualityProfile.pdf"), width = 8, height = 8, pointsize = 12)
for(i in ii) {
  message(paste0("[", i ,"/", length(Nova_sample_names), "] ", Nova_sample_names[i]))
  print(plotQualityProfile(fnFwds_Nova[i]) + ggtitle("Fwd"))
  print(plotQualityProfile(fnRevs_Nova[i]) + ggtitle("Rev"))
}
invisible(dev.off())

##Filter and Trim
filtFwds_Nova = file.path(Nova_filtered, basename(fnFwds_Nova))
filtRevs_Nova = file.path(Nova_filtered, basename(fnRevs_Nova))

out_Nova = filterAndTrim(fnFwds_Nova, filtFwds_Nova, fnRevs_Nova, filtRevs_Nova,
                    truncLen = c(140,140), minLen = 120, maxN = 0, truncQ = 2, maxEE =c(2,2),
                    rm.phix = TRUE, compress = TRUE, verbose = TRUE, multithread = TRUE)

out_Nova = as.data.frame(out_Nova)
rownames(out_Nova) = Nova_sample_names
head(out_Nova, 10)

##Learn the Error Rates
errFwd_Nova = learnErrors(filtFwds_Nova, multithread = TRUE)
errRev_Nova = learnErrors(filtRevs_Nova, multithread = TRUE)

pdf(paste0(Nova_images, "/plotErrors.pdf"), width = 10, height = 10, pointsize = 12)
plotErrors(errFwd_Nova, nominalQ = TRUE)
plotErrors(errRev_Nova, nominalQ = TRUE)
invisible(dev.off())

##Sample Inference
dadaFwds_Nova = dada(filtFwds_Nova, err = errFwd_Nova, pool = FALSE, multithread = TRUE)
dadaRevs_Nova = dada(filtRevs_Nova, err = errRev_Nova, pool = FALSE, multithread = TRUE)

##Merge Paired Reads
merged_Nova = mergePairs(dadaFwds_Nova, filtFwds_Nova, dadaRevs_Nova, filtRevs_Nova, verbose = TRUE)

##Construct the Sequence Table
seqtab_Nova <- makeSequenceTable(merged_Nova)

table(nchar(getSequences(seqtab_Nova)))

##Remove Chimeras
seqtab_nochim_Nova <- removeBimeraDenovo(seqtab_Nova, method="consensus", multithread=2, verbose=TRUE)
##Percent of sequences that are non-chimeric 
sum(seqtab_nochim_Nova)/sum(seqtab_Nova)

rownames(seqtab_nochim_Nova) <- Nova_sample_names

##Track Reads through the Pipeline
getN <- function(x) sum(getUniques(x))
track_Nova <- cbind(out_Nova, sapply(dadaFwds_Nova, getN), sapply(dadaRevs_Nova, getN), sapply(merged_Nova, getN), rowSums(seqtab_nochim_Nova))
colnames(track_Nova) <- c("input", "filtered", "denoisedF", "denoisedR", "merged","nonchim")
rownames(track_Nova) <- Nova_sample_names
head(track_Nova)

##Assign Taxonomy
seqs_Nova = getSequences(seqtab_nochim_Nova)
dbpath = "/16S_DB/"
ref_db = paste0(dbpath, "silva_nr99_v138.1_train_set.fa.gz")

taxonomy_tab_Nova <- assignTaxonomy(seqs_Nova, refFasta = ref_db)

Nova_metadata <- read.csv("https://raw.githubusercontent.com/reaset41/Infant-Urobiome/main/NovaSeq_meta.csv",header = TRUE,row.names = 1)

##Create Phyloseq Object
Nova_phyloseq_unfiltered <- phyloseq(otu_table(seqtab_nochim_Nova, taxa_are_rows=FALSE), 
               tax_table(taxonomy_tab_Nova), sample_data(Nova_metadata))


```


# Using Decontam to remove potential contaminant sequences 
We use the R package Decontam (citation: PMID 30558668) to identify sequences that are more prevalent in the negative controls (extraction blanks and PCR blanks) than the patient samples. First, using the known composition of the mock microbial community, we benchmarked Decontam performance to retain true sequence variants. Next, we applied the same threshold to patient samples and sampling controls to remove contaminants. 

## Applying Decontam to MiSeq mock communities to identify optimal Decontam threshold

```{r, eval=FALSE}
##Load requisite libraries
library("tidyverse")
library("decontam")
library("phyloseq")

##We would like to credit the following publication for guiding this code: Controlling for Contaminants in Low-Biomass 16S rRNA Gene Sequencing Experiments (PMID 31164452)
##https://lakarstens.github.io/ControllingContaminants16S/Analyses/ControllingContaminants16S_decontam.html

##Using phyloseq object created above 
MiSeq_phyloseq_unfiltered

##agglomerate ASVs with the same genus
MiSeq_phyloseq_unfiltered <- tax_glom(MiSeq_phyloseq_unfiltered, taxrank=rank_names(MiSeq_phyloseq_unfiltered)[6], NArm=TRUE, bad_empty=c(NA, "", " ", "\t"))

##Create subset of data limited to serial dilutions of the mock microbial community
mock_MiSeq <-subset_samples(MiSeq_phyloseq_unfiltered, !is.na(dilution))
sorted <- head(sort(colSums(otu_table(mock_MiSeq)), decreasing = TRUE), 8) #mock community includes 8 bacteria
mock_taxa <- cbind(as.data.frame(tax_table(mock_MiSeq)[names(sorted),]), Count = sorted)

##Create subset of data with only mock and blank data 
mock_blank_MiSeq <- subset_samples(MiSeq_phyloseq_unfiltered, !Type=="Patient")
##Normalize sample counts 
mock_blank_MiSeq <- transform_sample_counts(mock_blank_MiSeq,function(x) 100* x/sum(x))

##extract out original normalized ASV table
original_ASVs <- as.matrix(as.data.frame(mock_blank_MiSeq@otu_table))
##identify the original proportion of contaminants
contaminants_original <- rowSums(original_ASVs[,!colnames(original_ASVs) %in% rownames(mock_taxa)])
##identify the original proportion of mock community ASVs
mock_original <- rowSums(original_ASVs[,colnames(original_ASVs) %in% rownames(mock_taxa)])

##Evaluate different thresholds of Decontam
contam_mock_MiSeq <- isNotContaminant(mock_blank_MiSeq, neg = "is.neg", method = "prevalence", threshold = 0.3, normalize = FALSE, detailed = TRUE)

##create a phyloseq object without contaminants
mock_blank_MiSeq_true <- prune_taxa(contam_mock_MiSeq$not.contaminant, mock_blank_MiSeq)
##normalize the post-Decontam phyloseq object to 100
mock_blank_MiSeq_true_norm <- transform_sample_counts(mock_blank_MiSeq_true,function(x) 100* x/sum(x))

##subset out the ASV table of recovered ASVs
recovered_ASVs <- as.matrix(as.data.frame(mock_blank_MiSeq_true_norm@otu_table))

##create a subset of removed ASVs
removed_ASVs <- as.matrix(as.data.frame(mock_blank_MiSeq@otu_table))
removed_ASVs <- removed_ASVs[,!colnames(removed_ASVs) %in% colnames(recovered_ASVs)]

# Summarize results  
##proportion of contaminants removed (of all total contaminant ASVs)
contaminants_removed = (rowSums(removed_ASVs[,!colnames(removed_ASVs) %in% rownames(mock_taxa)])/ contaminants_original)
##proportion of mock removed (of all total mock ASVs)
mock_ASVs_removed = removed_ASVs[,(colnames(removed_ASVs) %in% rownames(mock_taxa))]/ mock_original

##total amount of contaminants remaining in new phyloseq object
contaminants_remaining =  rowSums(recovered_ASVs[,!colnames(recovered_ASVs) %in% rownames(mock_taxa)])

# Calculate classifications

##Percent of mock community ASVs correctly classified as mock community ASVs 
true_neg <- rowSums(recovered_ASVs[,colnames(recovered_ASVs) %in% rownames(mock_taxa)])
##Percent of mock community incorrectly classified as mock community ASVs
false_neg <- rowSums(recovered_ASVs[,!colnames(recovered_ASVs) %in% rownames(mock_taxa)])
##identify non-mock community ASVs correctly classified as not belonging to mock community
true_pos <- rowSums(removed_ASVs[,!colnames(removed_ASVs) %in% rownames(mock_taxa)])
# #identify mock community ASVs incorrectly classified as not belonging to mock community
false_pos <- rowSums(removed_ASVs[,colnames(removed_ASVs) %in% rownames(mock_taxa)])

sensitivity <- true_pos/(true_pos + false_neg)
specificty <- true_neg/(true_neg + false_pos)
accuracy <- (true_pos + true_neg) / (false_pos + true_pos + false_neg + true_neg)
prevalence <- (true_pos + false_neg) / (false_pos + true_pos + false_neg + true_neg)

##Following testing various thresholds, we settled on 0.3 as the ideal threshold for identifying contaminants and retaining true ASVs


```

## Applying Decontam to NovaSeq mock communities to identify optimal Decontam threshold

```{r, eval=FALSE}
##Load requisite libraries
library("tidyverse")
library("decontam")
library("phyloseq")

##Using phyloseq object created above 
Nova_phyloseq_unfiltered

##agglomerate ASVs with the same genus
Nova_phyloseq_unfiltered <- tax_glom(Nova_phyloseq_unfiltered, taxrank=rank_names(Nova_phyloseq_unfiltered)[6], NArm=TRUE, bad_empty=c(NA, "", " ", "\t"))

##Create subset of data limited to serial dilutions of the mock microbial community
mock_Nova <-subset_samples(Nova_phyloseq_unfiltered, !is.na(dilution))
sorted <- head(sort(colSums(otu_table(mock_Nova)), decreasing = TRUE), 8) #mock community includes 8 bacteria
mock_taxa <- cbind(as.data.frame(tax_table(mock_Nova)[names(sorted),]), Count = sorted)

##Create subset of data with only mock and blank data 
mock_blank_Nova <- subset_samples(Nova_phyloseq_unfiltered, !Type=="SamplingControl")
##Normalize sample counts 
mock_blank_Nova_norm <- transform_sample_counts(mock_blank_Nova_norm,function(x) 100* x/sum(x))

##extract out original normalized ASV table
original_ASVs <- as.matrix(as.data.frame(mock_blank_Nova_norm@otu_table))
##identify the original proportion of contaminants
contaminants_original <- rowSums(original_ASVs[,!colnames(original_ASVs) %in% rownames(mock_taxa)])
##identify the original proportion of mock community ASVs
mock_original <- rowSums(original_ASVs[,colnames(original_ASVs) %in% rownames(mock_taxa)])

##Evaluate different thresholds of Decontam
contam_mock_Nova <- isNotContaminant(mock_blank_Nova_norm, neg = "is.neg", method = "prevalence", threshold = 0.3, normalize = TRUE, detailed = TRUE)

##create a phyloseq object without contaminants
mock_blank_Nova_true <- prune_taxa(contam_mock_Nova$not.contaminant, mock_blank_Nova)
##normalize the post-Decontam phyloseq object to 100
mock_blank_Nova_true_norm <- transform_sample_counts(mock_blank_Nova_true,function(x) 100* x/sum(x))

##subset out the ASV table of recovered ASVs
recovered_ASVs <- as.matrix(as.data.frame(mock_blank_Nova_true_norm@otu_table))

##create a subset of removed ASVs
removed_ASVs <- as.matrix(as.data.frame(mock_blank_Nova@otu_table))
removed_ASVs <- removed_ASVs[,!colnames(removed_ASVs) %in% colnames(recovered_ASVs)]

# Summarize results  
##proportion of contaminants removed (of all total contaminant ASVs)
contaminants_removed = (rowSums(removed_ASVs[,!colnames(removed_ASVs) %in% rownames(mock_taxa)])/ contaminants_original)
##proportion of mock removed (of all total mock ASVs)
mock_ASVs_removed = removed_ASVs[,(colnames(removed_ASVs) %in% rownames(mock_taxa))]/ mock_original

##total amount of contaminants remaining in new phyloseq object
contaminants_remaining =  rowSums(recovered_ASVs[,!colnames(recovered_ASVs) %in% rownames(mock_taxa)])

# Calculate classifications

##Percent of mock community ASVs correctly classified as mock community ASVs 
true_neg <- rowSums(recovered_ASVs[,colnames(recovered_ASVs) %in% rownames(mock_taxa)])
##Percent of mock community incorrectly classified as mock community ASVs
false_neg <- rowSums(recovered_ASVs[,!colnames(recovered_ASVs) %in% rownames(mock_taxa)])
##identify non-mock community ASVs correctly classified as not belonging to mock community
true_pos <- rowSums(removed_ASVs[,!colnames(removed_ASVs) %in% rownames(mock_taxa)])
# #identify mock community ASVs incorrectly classified as not belonging to mock community
false_pos <- rowSums(removed_ASVs[,colnames(removed_ASVs) %in% rownames(mock_taxa)])

sensitivity <- true_pos/(true_pos + false_neg)
specificty <- true_neg/(true_neg + false_pos)
accuracy <- (true_pos + true_neg) / (false_pos + true_pos + false_neg + true_neg)
prevalence <- (true_pos + false_neg) / (false_pos + true_pos + false_neg + true_neg)

##Following testing various thresholds, we settled on 0.3 as the ideal threshold for identifying contaminants and retaining true ASVs


```

## Applying Decontam to MiSeq sequencing run to identify potential contaminants within patient urine samples 

```{r, eval=FALSE}
##Load requisite libraries
library("tidyverse")
library("decontam")
library("phyloseq")

##Using phyloseq object created above 
MiSeq_phyloseq_unfiltered

##agglomerate ASVs with the same genus
MiSeq_phyloseq_unfiltered <- tax_glom(MiSeq_phyloseq_unfiltered, taxrank=rank_names(MiSeq_phyloseq_unfiltered)[6], NArm=TRUE, bad_empty=c(NA, "", " ", "\t"))

##create subset of data excluding mock community samples
Patient_Blank_MiSeq <- subset_samples(MiSeq_phyloseq_unfiltered, !Type=="Mock")
##Normalize sample counts 
Patient_Blank_MiSeq <- transform_sample_counts(Patient_Blank_MiSeq,function(x) 100* x/sum(x))

##apply Decontam
contam_Patient <- isNotContaminant(Patient_Blank_MiSeq, neg = "is.neg", method = "prevalence",threshold = 0.3, normalize = FALSE, detailed = TRUE)


##create physloseq object without contaminants
physeq_true_MiSeq <- prune_taxa(contam_Patient$not.contaminant, Patient_Blank_MiSeq)
# normalize the post-Decontam phyloseq object to 100
physeq_true_MiSeq_norm <- transform_sample_counts(physeq_true_MiSeq,function(x) 100* x/sum(x))
##Subset to patient results only 
patient_true <- subset_samples(physeq_true_MiSeq_norm, Type=="Patient")

##remove chloroplast, mitochondrial, and unspecified taxa
patient_true_filtered <-subset_taxa(patient_true, Kingdom == "Bacteria" & !is.na(Phylum) & !is.na(Class) & Family != "Mitochondria" & Genus != "Mitochondria" & Order!="Chloroplast")

##remove ASVs <1% prevalence in entire patient dataset 
(0.01*sum(rowSums(otu_table(patient_true_filtered))))

filter_level <- colSums(otu_table(patient_true_filtered))>(0.01*sum(rowSums(otu_table(patient_true_filtered))))

patient_true_final <-prune_taxa(filter_level, patient_true_filtered)

patient_true_final

```

## Applying Decontam to NovaSeq sequencing run to identify potential contaminants within Sampling Controls 

```{r, eval=FALSE}
##Load requisite libraries
library("tidyverse")
library("decontam")
library("phyloseq")

##Using phyloseq object created above 
Nova_phyloseq_unfiltered

##agglomerate ASVs with the same genus
Nova_phyloseq_unfiltered <- tax_glom(Nova_phyloseq_unfiltered, taxrank=rank_names(Nova_phyloseq_unfiltered)[6], NArm=TRUE, bad_empty=c(NA, "", " ", "\t"))


##create subset of data excluding mock community samples
SamplingControl_Blank_Nova <- subset_samples(Nova_phyloseq_unfiltered, !Type=="Mock")
##Normalize sample counts 
SamplingControl_Blank_Nova <- transform_sample_counts(SamplingControl_Blank_Nova,function(x) 100* x/sum(x))

##apply Decontam
contam_SamplingControls <- isNotContaminant(SamplingControl_Blank_Nova, neg = "is.neg", method = "prevalence",threshold = 0.3, normalize = FALSE, detailed = TRUE)


##create physloseq object without contaminants
physeq_true_Nova <- prune_taxa(contam_SamplingControls$not.contaminant, SamplingControl_Blank_Nova)
# normalize the post-Decontam phyloseq object to 100
physeq_true_Nova_norm <- transform_sample_counts(physeq_true_Nova,function(x) 100* x/sum(x))
##Subset to sampling controls  only 
sampling_true <- subset_samples(physeq_true_Nova_norm, Type=="SamplingControl")

##remove chloroplast, mitochondrial, and unspecified taxa
sampling_true_filtered <-subset_taxa(sampling_true, Kingdom == "Bacteria" & !is.na(Phylum) & !is.na(Class) & Family != "Mitochondria" & Genus != "Mitochondria" & Order!="Chloroplast")



sampling_true_final <- sampling_true_filtered


```

# Figures 

## Figure 2A
```{r, eval=FALSE}
##Load requisite libraries
library(tidyverse)
library(phyloseq)
library(vegan)

merged_patient_sampling <- merge_phyloseq(patient_true_final,sampling_true_final)

dist <- phyloseq::distance(merged_patient_sampling, method = "bray")
ord <- ordinate(merged_patient_sampling, method="PCoA", distance = "bray")

Figure2A <- plot_ordination(merged_patient_sampling, ord, color = "Type") + 
  geom_point(size=3) + 
  stat_ellipse(aes(group=Type))+
  scale_color_manual(values = GroupsColors)+
  theme_bw()+ 
  theme(panel.grid = element_blank(),
        panel.border = element_rect(color="black",fill=NA,size=1.5),
        axis.text=element_text(size=13),
        axis.title=element_text(size=18))
        panel.background = element_rect(fill = "transparent"),
        plot.background = element_rect(fill = "transparent", color = NA),
        legend.background = element_rect(fill = "transparent"))


samples <- data.frame(sample_data(merged_patient_sampling))
adonis2(dist ~ Type, data = samples)

```

## Figure 2B

```{r, eval=FALSE}
##Load requisite libraries
library("tidyverse")
library("microbiome")
library("phyloseq")

sample_list <- c("H1", "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12", "H13", "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", "H24", "H25", "H26", "H27", "H28", "H29", "H30", "H31", "H32", "H33", "H34", "H35", "H36", "H37", "H38", "H39", "H40", "H41", "H42", "H43", "H44", "H45", "H46", "H47", "H48", "H49", "H50")

Figure2B <- comp_barplot(patient_true_final, tax_level = "Phylum", n_taxa = 5, bar_width = 0.7, sample_order=sample_list)+ coord_flip()+
  theme(axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=7),
        axis.title.x = element_text(size=15))


```

## Figure 2C

```{r, eval=FALSE}
##Load requisite libraries
library("tidyverse")
library("decontam")
library("phyloseq")
library("ggpubr")

alpha_stats <- estimate_richness(patient_true_final, measures = c("Chao1", "Shannon"))
GroupsColors <- c("#6db6ff", "#009292")
comparisons_delivery <- list( c("Vaginal", "Caesarean"))
comparisons_abx <- list( c("Yes", "No"))
symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns"))

alpha_stats <- estimate_richness(patient_true_final, measures = c("Chao1", "Shannon"))
GroupsColors <- c("#6db6ff", "#009292")
comparisons_delivery <- list( c("Vaginal", "Caesarean"))
comparisons_abx <- list( c("Yes", "No"))
symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns"))

Delivery_Stat <- plot_richness(patient_true_final, x="delivery", measures=c("Chao1", "Shannon"), color = "delivery")+
  geom_boxplot(alpha=0.6, size=1.5)+ 
  geom_point(size=4)+
  theme_bw()+
  theme(axis.text = element_text(size = 16))+
  scale_color_manual(values = GroupsColors)+
  theme(legend.position="none", axis.text.x=element_text(angle=0,hjust=0.5,vjust=1,size=8))+
  stat_compare_means(method = "wilcox.test", comparisons = comparisons_delivery, label = "p.signif", symnum.args = symnum.args, size=6)

Abx_Stat <- plot_richness(patient.true, x="abx", measures=c("Chao1", "Shannon"), color = "abx")+
  geom_boxplot(alpha=0.6, size=1.5)+ 
  geom_point(size=4)+
  theme_bw()+
  theme(axis.text = element_text(size = 16))+
  scale_color_manual(values = GroupsColors)+
  theme(legend.position="none", axis.text.x=element_text(angle=0,hjust=0.5,vjust=1,size=9))+
  stat_compare_means(method = "wilcox.test", comparisons = comparisons_abx, label = "p.signif", symnum.args = symnum.args, size=6)



```

## Figure 2D

```{r, eval=FALSE}
##Load requisite libraries
library("tidyverse")
library("phyloseq")
library("ggpubr")

lacto <- subset_taxa(patient_comp, Family=="Lactobacillaceae")

df_lacto <- as.data.frame(colSums(abundances(lacto)))
df_lacto$Delivery <- MiSeq_metadata$delivery
colnames(df_lacto) <- c("Lacto_Abund","Delivery")

symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns"))
colorpal <- c("#E69F00", "#56B4E9")
comparisons_delivery <- list( c("Vaginal", "Caesarean"))

ggplot(df_lacto, aes(x=Delivery, y=Lacto_Abund, group=Delivery))+
  geom_boxplot(alpha=0.6, size=1, aes(col=Delivery))+ 
  geom_point(size=4, aes(col=Delivery))+
  scale_color_manual(values=GroupsColors)+
  theme_bw()+
  theme(axis.text = element_text(size = 12))+
  stat_compare_means(method = "wilcox.test", comparisons = comparisons_delivery, label = "p.signif", symnum.args = symnum.args, size=6)


```

## Figure 3

```{r, eval=FALSE}
library(tidyverse)
library(reshape2)

concordance <- read.csv("https://raw.githubusercontent.com/reaset41/Infant-Urobiome/main/concordance_Figure3.csv",header = TRUE)

melt_concord <- melt(concordance)

##0 = not detected
##1 = EQUC only
##2 = 16S rRNA only
##3 = Both methods
col1 <- c("0", "1", "2", "3")

Concord <- ggplot(melt_concord, aes(variable, Family, fill=factor(value))) + 
  geom_tile(color = "white",linewidth = 0.25,linetype = 1)+
  coord_fixed()+
  scale_fill_manual(values = c("grey", "#800000", "#ADD8E6", "#000080"))+
  xlab("Patient Number")+
  theme(
    axis.text.x   = element_text(colour = "black", angle=45, size=5),
    axis.title.x = element_text(colour = "black", size = rel(0.5)),
    axis.text.y = element_text(colour = "black", size=5))

```


# R session information

```{r}

sessionInfo()

```

